---
title: Trademarks from the Oregon Secretary of State
author: RWW
date: '2022-01-26'
slug: []
categories:
  - ggplot
  - GSM622
tags: []
---

```{r setup, include=FALSE}
library(tidyverse)
library(RSocrata)
```

# Oregon's Open Data Portal

The Oregon Open Data Portal has a bunch of easily accessible public use data.  One of the first listings today was trademarks issued by the Oregon Secretary of State.  Let's grab that data.

```{r}
OTM <- read.socrata('https://data.oregon.gov/resource/ny3n-dx3v.json')
head(OTM)
```

I want to plot the copyrights by state of holder.  Let's look at that variable.

```{r}
library(janitor) # for tabyl
OTM %>% tabyl(state)
```

There are a few to clean up.  I will use recode to do that for many of them and split out the junk text from those that have some.

```{r}
( OTM %>% 
  mutate(state = recode(state, 
                        `0R` = "OR", 
                        `O` = "OR", 
                        `OROR` = "OR", 
                        `ORQ` = "OR")) %>% 
  mutate(state2 = str_split(state, " ", simplify=TRUE)) %>% 
  filter(state2!="NA") %>% 
  janitor::tabyl(state2) -> ORTM.Data )
```

That should get me what I need.  Now to plot it, sorted from most to least.

```{r}
ORTM.Data %>% ggplot() + 
  aes(x=fct_reorder(state2, n), y=n, fill=state2) + 
  geom_col() + 
  guides(fill="none") +
  coord_flip() + 
  labs(title="Trademarks Issued by OR SOS", x="State", y='Trademarks', caption="https://data.oregon.gov/resource/ny3n-dx3v.json")
```

