---
title: "TT: Beyoncé and Taylor Swift Lyrics"
authors: ["RWW"]
date: '2020-10-19'
slug: beyonce-words
categories:
  - tidyTuesday
  - dataviz
tags:
  - tidyTuesday
subtitle: ''
summary: ''
lastmod: '2020-10-19T02:11:17-07:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
library(here)
```

## tidyTuesday: Beyoncé and Taylor Swift Lyrics

<img src="https://camo.githubusercontent.com/9fd20607eb6961861f803167a3d1f17a3724d360af97051e556eeaeb516ab14e/68747470733a2f2f616b6e732d696d616765732e656f6e6c696e652e636f6d2f656f6c5f696d616765732f456e746972655f536974652f323031393732312f72735f31303234783735392d3139303832313132353131322d313032342e7461796c6f722d73776966742d6265796f6e63652d323030392d6d74762d766d61732e63742e3038323131392e6a70673f6669743d61726f756e64253743313032343a6175746f266f75747075742d7175616c6974793d39302663726f703d313032343a6175746f3b63656e7465722c746f70" alt="tidytuesday header photo of beyonce and taylor swift" width="200"/>

`tidyTuesday` for the final week of September 2020 is based on the music of  Beyoncé and Taylor Swift.  To be honest, I do not know either artist well so I will pick Beyoncé and look at her lyrics.  The raw data are organized as a rather typical text file though there is some underlying tidyness to the rows and songs as embedded data to work with.  I will not work much with it but it is there to exploit.  First, I load the data.

```{r}
beyonce_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')
str(beyonce_lyrics)
```

## Beyonce's Favorite Lyric Words

What words does Beyoncé use the most?  I will deploy `tidytext` tools to unnest the word tokens, get rid of stop words, and create a barplot from the table of frequencies.

```{r}
beyonce_lyrics %>% 
  unnest_tokens(word, line) %>%   # Parse the lyric lines to tidy: one word per row.
  anti_join(., stop_words) %>%    # Remove the stop words
  group_by(word) %>%              # Group them by word
  summarise(Count = n()) %>%      # How common is the word?
  top_n(25) %>%                   # Keep the top 25
  ggplot() + 
  aes(x=fct_reorder(word, Count), y=Count) + 
  geom_col() + 
  coord_flip() + 
  labs(x="Word", title="Beyoncé's Top Words in Lyrics")
```

## Beyoncé's Cloud

```{r BC1}
library(wordcloud2)
beyonce_lyrics %>% 
  unnest_tokens(word, line) %>%   # Parse the lyric lines to tidy: one word per row.
  anti_join(., stop_words) %>%    # Remove the stop words
  group_by(word) %>%              # Group them by word
  summarise(Count = n()) %>%      # How common is the word?
  wordcloud2::wordcloud2(size = 0.7, shuffle = TRUE) -> WC111
# htmlwidgets::saveWidget(widgetframe::frameWidget(WC111), file='widgets/wcbey.html')
# MyWC
# widgetframe::frameWidget(WC111)
WC111
```

I will finish with the sentence rendition of the graphic.

```{r BC2}
library(wordcloud2)
MyDat <- beyonce_lyrics %>% 
  unnest_tokens(word, line) %>%   # Parse the lyric lines to tidy: one word per row.
  anti_join(., stop_words) %>%    # Remove the stop words
  group_by(word) %>%              # Group them by word
  summarise(freq = n()) %>%      # How common is the word?
  arrange(desc(freq))
MyWC <- wordcloud2::letterCloud(MyDat, word="Most Popular \n Words in \n Beyonce \n Lyrics", wordSize = 1.75, backgroundColor="lightblue", color='random-dark', size=0.8)
# htmlwidgets::saveWidget(widgetframe::frameWidget(MyWC), file='widgets/beyonce.html')
# MyWC
# widgetframe::frameWidget(MyWC)
```

Something about this does not seem to work now as I redesign the website.  It did at one time.

