---
title: Top 5 Team Player Valuations
author: RWW
date: '2022-01-28'
slug: []
categories:
  - dataviz
  - ggplot
  - tidyverse
tags:
  - tidyverse
---

## Data

There is a ton of data in the `worldfootballR` package; the github for the package is [linked here](https://github.com/JaseZiv/worldfootballR).  This particular analysis uses data from `transfermarkt`.  I am going to grab player market values for the top five European leagues.

```{r, message=FALSE, eval=TRUE, cache=TRUE}
library(tidyverse); library(here)
library(worldfootballR)
options(scipen=9)
Big_5_player_values <- get_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"), start_year = 2021)
save(Big_5_player_values, file=here("content/post/2022-01-28-top-5-team-player-valuations/data","Big5.RData"))
```

```{r, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
library(tidyverse); library(here)
library(worldfootballR)
options(scipen=9)
load(here("content/post/2022-01-28-top-5-team-player-valuations/data","Big5.RData"))
```


## A Money Table Function

I want to repeat the same basic plot five times, only selecting a particular league in each iteration.  That is a perfect case to write a function and then invoke it for each league.  I need a function that only takes one argument, the `comp_name` that I want to pull out and plot.  The competition names will then be the inputs.

```{r}
Plot.Money.Table <- function(value) {
  Big_5_player_values %>% 
    group_by(comp_name, squad) %>% 
    summarise(Team.Value = sum(player_market_value_euro, na.rm=TRUE)/100000000) %>%
    ungroup() %>%
    filter(comp_name==value) %>% 
    ggplot() + 
    aes(x=fct_reorder(squad, Team.Value), y=Team.Value) + 
    geom_col() + 
    coord_flip() + 
    labs(x="Team", y="Total Value of Players in 100M Euros", title=paste0(value, " Money Table", sep=""), caption="Data from transfermarkt") 
}
```

## The English Money Table

```{r}
Plot.Money.Table("Premier League")
```

## The Spanish Money Table

```{r}
Plot.Money.Table("LaLiga")
```

## The Italian Money Table

```{r}
Plot.Money.Table("Serie A")
```

# The Germany Money Table

```{r}
Plot.Money.Table("Bundesliga")
```

# The French Money Table

```{r}
Plot.Money.Table("Ligue 1")
```
