---
title: National Weather Service Portland, Part II
author: RWW
date: '2022-03-18'
slug: nws-monthly-part-ii
categories:
  - ggplot
  - tidyverse
  - time series
tags:
  - time series
  - tidyverse
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(tidyverse)
library(lubridate)
library(hrbrthemes)
library(kableExtra)
library(magrittr)
library(fpp3)
```

## Loading NWS Data

I first will try to load it without any intervention to see what it looks like.  As we will see, it is quite messy in a few easy ways and a few that are a bit more tricky.

```{r}
NWS <- read.csv(url("https://www.weather.gov/source/pqr/climate/webdata/Portland_dailyclimatedata.csv"))
head(NWS, 10) %>% kable() %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "500px")
```

The column names are stored in the seventh row; to properly import this.  In addition, there are two missing value codes: `M` and `-` that will have to be accounted for.  I will use `skip` to skip the first 6 rows and declare two distinct values to be encoded as missing.  Let's see what we get.

```{r}
NWS <- read.csv(url("https://www.weather.gov/source/pqr/climate/webdata/Portland_dailyclimatedata.csv"), skip=6, na.strings = c("M","-"))
head(NWS, 10)
```

Two other things are of note.  The first one is that R really doesn't like columns to be named as numbers so we have an `X` in front of the numeric days.  The second is that the column denoting which variable the rows represent is now `X`.  Let me rename `X` to be `Variable`.


```{r}
NWS <- read.csv(url("https://www.weather.gov/source/pqr/climate/webdata/Portland_dailyclimatedata.csv"), skip=6, na.strings = c("M","-")) %>% 
  rename(Variable = X)
str(NWS)
```

It is disappointing that everything is stored as `character` type.  That will prove advantageous in one respect because there is some `/A` garbage embedded in two of the variables (SN and PR).  Here, I will ask R to find all columns that are stored as character and ask it to remove the string.

```{r}
NWS <- NWS %>% mutate(across(where(is.character), ~str_remove(.x, "/A")))
```

Now, we will have to fix the values `T` in the precipitation and snow variables [which are currently stored in repeated rows].  Nevertheless, this should give me what I need to create the monthly data.

### Daily Data

The daily data will necessarily not involve the column of Totals/Averages that we used for the monthly data so let us eliminate it.

```{r}
NWS.Daily <- NWS %>% select(-AVG.or.Total)
```

Now I want to rename the columns with names X1, X2, ..., X31 to `Day.1` to `Day.31` for clarity.  It is largely inconsequential, it would work on the X's but I prefer nicely labeled intermediate steps.

```{r}
names(NWS.Daily) <- c("YR","MO","Variable",paste0("Day.",1:31))
```

The next step is to tidy the data.  First, let me use `pivot_longer` on every column that starts with `Day.` putting the variable names in `Day` and variable values in `value`.

```{r}
NWS.Daily.Base <- NWS.Daily %>% 
  pivot_longer(., cols=starts_with("Day."), names_to = "Day", values_to = "value")
head(NWS.Daily.Base)
```

Now I want to turn the days into numbers [they are character above] and then use `pivot_wider` to get the four variables into unique columns, recode trace [T] where they exist to numbers that are half the size of the smallest values, turn them into numbers, and create a date.

```{r}
NWS.Daily.Base %<>%  mutate(Day = str_remove(Day, "Day.")) %>%  
  pivot_wider(., names_from = "Variable", values_from = "value") 
NWS.Daily <- NWS.Daily.Base %>% mutate(PR = recode(PR, T = "O.005"), SN = recode(SN, T = "O.005")) %>% 
  mutate(TX = as.numeric(TX), TN = as.numeric(TN), PR = as.numeric(PR), SN = as.numeric(SN), 
         date = as.Date(paste(MO,Day,YR,sep="-"), format="%m-%d-%Y")
         )
NWS.Daily.Clean <- NWS.Daily %>% filter(!is.na(date))
head(NWS.Daily.Clean)
```

This is exactly what I needed.

# Some Plots


# By Day?

```{r}
NWS.Daily.Clean %>% ggplot() + aes(x=date, y=PR) + geom_point(alpha=0.1) + theme_ipsum_rc() + labs(title="Daily Precipitation")
```

This is really pretty terrible.  This is going to need a good bit of work.  My first go is going to be to create a moving average that can smooth out the look.  I will use a 14-day moving average.

```{r}
NWS.Daily.Clean %>% 
  arrange(date) %>% 
  mutate(Rolling.Average = zoo::rollmean(PR, 7, fill=NA)) %>%
  ggplot(., aes(x=date, y=PR)) + geom_point(alpha=0.05, size=0.5) + geom_line(aes(x=date, y=Rolling.Average), inherit.aes=FALSE, color="red") + theme_ipsum_rc() + labs(title="Daily Precipitation")
```


## Time Series Plots

### Daily

```{r}
NWS.Daily.Clean <- NWS.Daily.Clean %>% as_tsibble(., index=date)
NWS.Daily.Clean %>% filter(date > "2010-01-01") %>% autoplot(TX) + labs(title="Daily High Temperatures in Portland, Oregon", caption = "Data from NWS", x = "Date", y="High Temperature (deg F)")
```

## Seasonal Plots

### Daily

```{r}
NWS.Daily.Clean %>% filter(date > "2015-01-01") %>% gg_season(TX, labels = "both")
```

## Subseries Plots

### Daily

Daily subseries plots are a mess because there are 31 and I am not sure that there would be much instructive anyway as daily variation in temperature is quite noisy.

```{r}
NWS.Daily.Clean %>% gg_subseries(TX, period="weeks")
```

# Autocorrelation

To what degree are observations separated by k time periods correlated.

$$r_{k} = \frac{\sum\limits_{t=k+1}^T (y_{t}-\bar{y})(y_{t-k}-\bar{y})}
 {\sum\limits_{t=1}^T (y_{t}-\bar{y})^2}$$
where $T$ is the length of the time series. The autocorrelation coefficients make up the autocorrelation function or ACF.

The autocorrelation coefficients for the monthly high temperatures can be computed using the ACF() function.


### Daily: ACF

```{r}
NWS.Daily.Clean %>% ACF(TX) %>% autoplot()
```


