---
title: Cleaning Messy National Weather Service Data for Portland
author: RWW
date: '2022-03-19'
slug: nws-monthly-joins
categories:
  - ggplot
  - tidyverse
  - time series
tags:
  - time series
  - tidyverse
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(tidyverse)
library(lubridate)
library(hrbrthemes)
library(kableExtra)
library(magrittr)
library(fpp3)
```

## Loading NWS Data

I first will try to load it without any intervention to see what it looks like.  As we will see, it is quite messy in a few easy ways and a few that are a bit more tricky.

```{r}
NWS <- read.csv(url("https://www.weather.gov/source/pqr/climate/webdata/Portland_dailyclimatedata.csv"))
head(NWS, 10) %>% kable() %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "500px")
```

The column names are stored in the seventh row; to properly import this.  In addition, there are two missing value codes: `M` and `-` that will have to be accounted for.  I will use `skip` to skip the first 6 rows and declare two distinct values to be encoded as missing.  Let's see what we get.

```{r}
NWS <- read.csv(url("https://www.weather.gov/source/pqr/climate/webdata/Portland_dailyclimatedata.csv"), skip=6, na.strings = c("M","-"))
head(NWS, 10)
```

Two other things are of note.  The first one is that R really doesn't like columns to be named as numbers so we have an `X` in front of the numeric days.  The second is that the column denoting which variable the rows represent is now `X`.  Let me rename `X` to be `Variable`.


```{r}
NWS <- read.csv(url("https://www.weather.gov/source/pqr/climate/webdata/Portland_dailyclimatedata.csv"), skip=6, na.strings = c("M","-")) %>% 
  rename(Variable = X)
str(NWS)
```

It is disappointing that everything is stored as `character` type.  That will prove advantageous in one respect because there is some `/A` garbage embedded in two of the variables (SN and PR).  Here, I will ask R to find all columns that are stored as character and ask it to remove the string.

```{r}
NWS <- NWS %>% mutate(across(where(is.character), ~str_remove(.x, "/A")))
```

Now, we will have to fix the values `T` in the precipitation and snow variables [which are currently stored in repeated rows].  Nevertheless, this should give me what I need to create the monthly data.

### Monthly Data

Now I will have to facilitate type conversions; the data are currently character instead of numeric.  Looking at the data, I think it is reasonable, though not with total certainty, that they should take values bigger than 0 but smaller than the smallest measured values for that `Variable`; in this case, the smallest values that I note are 0.01.  I will (somewhat) arbitrarily set them to 0.005.  With that in mind, let's see if we can create the monthly data by adjusting those values and selecting only the year, the month. the variable, and `AVG.or.Total` -- the last column that contains the average high/low and the total snowfall and precipitation for that month of that year.  Finally, let me change the type of that last column to numeric.

```{r}
# Now to create a Monthly time series.  The first step is to select the four columns 
NWS.Monthly.Base <- NWS %>% 
  select(YR, MO, Variable, AVG.or.Total) %>% 
  mutate(AVG.or.Total = recode(AVG.or.Total, T = "O.005", `T/A` = "O.005")) %>% # Replace the trace amounts with numbers as they are not missing nor are they zero.
  mutate(AVG.or.Total = stringr::str_replace_all(AVG.or.Total, "/A", "")) %>%
  mutate(AVG.or.Total = as.numeric(AVG.or.Total)) # Turn it into numeric
summary(NWS.Monthly.Base)  # Show a summary
```
That seems to work.  Two more steps to get a clean monthly dataset.  The last step is to create four variables named TX, TN, PR, and SN as unique data.frames.  To make merging them by a common index easier, I will use paste to create a character variable of the data and then turn it into a Date: `Month.YR`.

```{r}
TX <- NWS.Monthly.Base %>% filter(Variable == "TX") %>% 
    mutate(Month.Yr = as.Date(paste0(MO,"-14-",YR), format = "%m-%d-%Y"),
           TX = AVG.or.Total) %>%
    select(-c(Variable,AVG.or.Total))
TN <- NWS.Monthly.Base %>% filter(Variable == "TN") %>% 
    mutate(Month.Yr = as.Date(paste0(MO,"-14-",YR), format = "%m-%d-%Y"),
           TN = AVG.or.Total) %>% 
    select(-c(Variable,AVG.or.Total))
PR <- NWS.Monthly.Base %>% filter(Variable == "PR") %>% 
    mutate(Month.Yr = as.Date(paste0(MO,"-14-",YR), format = "%m-%d-%Y"),
           PR = AVG.or.Total) %>%
    select(-c(Variable,AVG.or.Total))
SN <- NWS.Monthly.Base %>% filter(Variable == "SN") %>% 
    mutate(Month.Yr = as.Date(paste0(MO,"-14-",YR), format = "%m-%d-%Y"),
           SN = AVG.or.Total) %>%
    select(-c(Variable,AVG.or.Total))
```

The last step is to merge the clean individual series together; I will use `left_join`.

```{r}
NWS.Monthly.Clean <- left_join(TX,TN) %>% left_join(.,PR) %>% left_join(.,SN)
```

Woo hoo!  One down and one to go.  But, as we should've noticed before, their totals aren't actually great.

## Maximum Temperature

```{r}
NWS.Monthly.Clean %>% ggplot() + aes(x=Month.Yr, y=TX) + geom_line()
```

## Minimum Temperature

```{r}
NWS.Monthly.Clean %>% ggplot() + aes(x=Month.Yr, y=TN) + geom_line()
```

## Snow

```{r}
NWS.Monthly.Clean %>% ggplot() + aes(x=Month.Yr, y=SN) + geom_line()
```

## Precipitation

```{r}
NWS.Monthly.Clean %>% ggplot() + aes(x=Month.Yr, y=PR) + geom_line()
```
