---
title: Inflation Expectations
author: RWW
date: '2022-03-19'
slug: inflation-expectations
categories:
  - dataviz
  - ggplot
  - time series
  - tidyverse
tags:
  - time series
  - tidyverse
---

The Federal Reserve Bank of New York provides interesting data on inflation expectations.  I was first interested in this because it appears as though the survey respondents have consistently overestimated inflation expectations over the last few years.  But now, in a time of heightened concern about inflation, it is worthwhile to revisit the data.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate); library(tsibble)
library(readxl); library(magrittr)
library(kableExtra)
url <- "https://www.newyorkfed.org/medialibrary/interactives/sce/sce/downloads/data/frbny-sce-data.xlsx"
destfile <- "frbny_sce_data.xlsx"
curl::curl_download(url, destfile)
Inflation.Expectations <- read_excel(destfile, sheet=4, skip=3) 
Inflation.Expectations %>% 
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

Now let me format the dates properly to treat these data as a time series.

```{r}
Inflation.Expectations %<>%
  rename(date = 1) %>% 
  mutate(date = yearmonth(parse_date_time(date, orders = "%Y%m")))
```

That's pretty much all that is required to visualize them.  Let's have a look.  I will use the `fpp3` library and `autoplot()`.

```{r}
library(fpp3)
Inflation.Expectations %>% 
  as_tsibble(index=date) %>% 
  autoplot(`Median three-year ahead expected inflation rate`) + 
  geom_line(aes(y=Inflation.Expectations$`Median one-year ahead expected inflation rate`, color="red")) +
  hrbrthemes::theme_ipsum() + guides(color = "none")
```

Now let me pretty the plot.

```{r}
Inflation.Expectations %>% 
  select(1:3) %>% 
  pivot_longer(c(2:3)) %>%
  mutate(Variable = name) %>%
  ggplot(aes(x=date, y=value, color=Variable)) +
  geom_line(size=2) +
  scale_color_viridis_d(option="C") +
  hrbrthemes::theme_ft_rc() +
  theme(legend.position = "bottom") +
  labs(y="Inflation Expectations", 
       color="",
       title="New York Fed Inflation Expectations",
       caption = "data: https://www.newyorkfed.org/medialibrary/interactives/sce/sce/downloads/data/frbny-sce-data.xlsx")
```


