---
title: Mapping Voters
author: RWW
date: '2022-04-20'
slug: []
categories:
  - Maps
tags: []
---


## The Data

Robert Husseman [@RHusseman on Twitter](https://twitter.com/RHusseman/) is running to represent Oregon District 21 in the Oregon House of Representatives. He has a platform that focuses on housing and livability in the state by making sensible allocations of public resources to public problems. He has an MBA; he was my student and his head and heart are very much in the right place to serve Oregon well.

His campaign has a voter list for the Democratic voters in Oregon 21. He shared access to this. Let me load them up from a saved Excel spreadsheet derived from a Google sheet.

```{r, include=FALSE}
library(readxl)
RHCamp <- read_excel("~/Desktop/RHCamp.xlsx", 
    col_types = c("numeric", "text", "text", 
        "text", "text", "text", "text", "date", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "numeric", "numeric", 
        "text", "text", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric"))
library(tidyverse)
```

## Geocoding

I ran across a package called `tidygeocoder` to do the heavy lifting. Here is what I have to work with.

```{r}
library(tidygeocoder)
names(RHCamp)
```

## A Great Thing about R

There is [a vignette explaining the usage.](https://cran.r-project.org/web/packages/tidygeocoder/vignettes/tidygeocoder.html)

My use case is pretty easy though there is a lot of data. I want to try this with just the addresses. I have loaded the `tidyverse` and now I want to peel off just the data that I need in order to geocode it, and then process it with the Census geocoder. This results in 9443 queries.

```{r, eval=FALSE}
GeoMe <- RHCamp %>% select(VOTER_ID, RES_ADDRESS_1, CITY, STATE) %>% geocode(street = RES_ADDRESS_1, city=CITY, state=STATE, method="census")
NewData <- GeoMe %>% left_join(., RHCamp)
save(NewData, file="~/Desktop/NewDataRH.RData")
```

This takes quite a long time to run -- 667.1 seconds. 

When it is complete, I will want to figure out how to map them in a way that is best fit for assisting a campaign. This seems like a good use for `leaflet`. Rather than taking 10 minutes or so to compile each time, I saved them off. I am loading them from a local file in the first line.

## Leaflet Mapping

```{r}
load("~/Desktop/NewDataRH.RData")
```

The `leaflet` package uses very powerful mapping tools that will be perfect for this use case because the maps have easy zoom features.

```{r}
library(leaflet)
NewData %>% mutate(labMy = paste0(FIRST_NAME,LAST_NAME,RES_ADDRESS_1,sep="\n")) %>%
  leaflet(data = .) %>% 
  addTiles() %>%
  addCircleMarkers(~long, ~lat, label = ~labMy, fillOpacity = 0.1, stroke=FALSE, radius=1.6)
```

That works but it isn't quite right.  Turns out that I will have to mess with html to get the labels quite right.

So, of course, I went to `stackoverflow`.

## Building Better Labels

[I found a post on how to do this.](https://stackoverflow.com/questions/43144596/r-and-leaflet-how-to-arrange-label-text-across-multiple-lines)

The key piece of code seems to be:

```{r}
labs <- lapply(seq(nrow(NewData)), function(i) {
  paste0( '<p>', NewData[i,"RES_ADDRESS_1"],'</p>' )
})
NewData %>%
  leaflet(data = .) %>% 
  addTiles() %>%
  addCircleMarkers(~long, ~lat, label = lapply(labs, htmltools::HTML), fillOpacity = 0.15, stroke=FALSE, radius=1.6)
```
